---
title: "lab 9"
author: "Priscilla Hu"
date: "11/18/2019"
output: pdf_document
---
## Lab 09: Clustering

## 1. K-Means Clustering

```{r}
Euc_dist <- function(x,y){
 return(sqrt((x-y)^2)) 
}

L1_norm <- function(x,y){
  return(sum(abs(x-y))) 
}

my_kmeans <- function(x, k, d){
  n <- nrow(x)
  init <- sample(1:n, k)
  centroids <- x[init,]
  assignment <- apply(x, 1, function(obs) {
    distances <- apply(centroids, 1, function(centroid){
      d(obs,centroid)
    })
  which.min(distances)
  }) 
 repeat {
    prev <- assignment
    mat_list <- split(x,assignment)
    centroids <- t(sapply(mat_list,colMeans))
    assignment <- apply(x, 1, function(obs) {
      distances <- apply(centroids, 1,function(centroid) {
        d(obs,centroid)
    })
  which.min(distances)
  })
  if (all(assignment == prev)){
    break
   }     
 }

  wss <- sapply(split(x,assignment), function(x_sub) {
    sum(apply(x_sub, 2, function(col){ sum((col - mean(col))^2)} ))
  })
  
  tss <- sum(apply(x, 2, function(col){ sum((col - mean(col))^2)} ))
  bss_over_tss <- (tss - sum(wss)) / tss
  return(list(cluster_sizes = as.vector(table(assignment)),cluster_means = centroids, clustering_vector = assignment, wss_cluster = wss, bss_over_tss = bss_over_tss))
}
```

```{r}
my_kmeans(iris[1:4],3,L1_norm)
kmeans(iris[1:4],3)
```


Conceptual Questions 
1. Suppose I perform PCA on my dataset to get U=XV, where U is the PC-transformed data matrix, X is the original data matrix, and V is the matrix of loadings. Using the Euclidean distance function, I then run k-means clustering on both U and X with everything equal (initial points, order of looping). Will the clustering solutions be the same? Why or why not? 

Ans:Use all the components and it just roatate our data.


2. Now I want to do dimension reduction and only keep the first L principal components before repeating Step 1. How will U change? Will the clustering solution stay the same? 

Ans:No. 


3. Now suppose I repeat Step 1 but with the L1 distance. Will the clustering solutions be the same? Why or why not? 

Ans:No. l1 dist is changed.


4. (Optional) Run some experiments to test your answers. Are the clustering solutions improved?




## 2. (Agglomerative) Heirarchical Clustering

```{r}
hc.complete <-hclust(dist(iris[, 1:4]), method="complete") 
plot(hc.complete, main="Complete Linkage ", xlab="", sub="",cex=.9)
```

```{r}
hc.single<-hclust(dist(iris[, 1:4]), method="single")
plot(hc.single, main="Single Linkage ", xlab="", sub="",cex=.9)
```

```{r}
hc.average <- hclust(dist(iris[, 1:4]), method="average")
plot(hc.average, main="Average Linkage ", xlab="", sub="",cex=.9)
```

```{r}
cutree(hc.complete,3)
```

Most of them are the same. It may vary in number but basically it’s fine.